---
title: "Fargo_Crime_2017"
output: html_notebook
---

```{r, warning = F}
library(tidyverse)
library(ggmap)
library(lubridate)
library(ggthemes)
```


#### Upload Data

Try uploading test of the full data, not all lat/lon included
```{r}
full <- read.csv("crime.csv", header = T)
```

```{r}
str(full)
```
#### Data Wrangling

Convert "Date.Time" field into actual date field
```{r}
full$Date.Time <- lubridate::mdy(full$Date.Time)
```


Convert "Time" field into actual time field
```{r}
library(lubridate)

full$Time <- lubridate::mdy_hms(full$Time)

```


```{r}
str(full)
```


Drop the NA rows
```{r}
full <- tidyr::drop_na(data = full)
```

```{r}
str(full)
```


```{r}
head(full)
```

Recode the Call.Type to smaller groups
```{r}
full.recode <- dplyr::mutate(full, Category = recode(full$Call.Type,   # make new categories
                                                       "Abuse" = "Violent",
                                                       "Assault" = "Violent",
                                                       "Bicycle - Theft" = "Property",
                                                       "Bomb Threat" = "Violent",
                                                       "Burglary" = "Property",
                                                       "Death Notification" = "Violent",
                                                       "Domestic" = "Violent",
                                                       "Graffiti" = "Property",
                                                       "Gun Shots Heard" = "Violent",
                                                       "Harassment" = "Violent",
                                                       "Impaired Driver" = "Drugs/Alcohol",
                                                       "Impaired Person" = "Drugs/Alcohol",
                                                       "Kidnapping" = "Violent",
                                                       "Loud Party/People" = "Drugs/Alcohol",
                                                       "Overdose - Poisoning" = "Drugs/Alcohol",
                                                       "Prowler/Peeper" = "Violent",
                                                       "Robbery" = "Property", 
                                                       "Sexual Assault" = "Violent",
                                                       "Stab/Gunshot/Penetrating" = "Violent",
                                                       "Stolen Vehicle" = "Property",
                                                       "Theft" = "Property",
                                                       "Vandalism" = "Property",
                                                       "Vehicle Break In" = "Property"))


    ```



#### Graphs Begin

## 1. Crime category by time of day
```{r}

ggplot(full.recode, aes(x=Time)) + geom_density(aes(fill=as.factor(Category), colour=Category, y=..density..), stat="density", position="identity", alpha=0.4) + theme_wsj(base_size = 8) + theme(text=element_text(family="sans", face="plain", color="#000000", size=8, hjust=0.5, vjust=0.5)) + guides(fill=guide_legend(title="Category"), colour=guide_legend(title="Category"))  + ylab("density") + xlab("Time of Day") + ggtitle("") +
scale_x_datetime(labels = date_format("%H:%M")) + theme(axis.text.y=element_blank()) 
ggsave(filename = "plot1.png")
```










Plot Calendar Heatmap
using ggTimeSeries

Wrangle data into occurences by day
```{r}
library(data.table)
calendar.data <- dplyr::add_count(full.recode, Date.Time)
calendar.data <- dplyr::arrange(calendar.data, Date.Time)

calendar.data <- calendar.data %>% 
  group_by(Date.Time) %>% 
  summarise(avg = mean(n))

# need to coerce to data.table
calendar.data <- data.table::data.table(Date = calendar.data$Date.Time,
                                        Crimes = calendar.data$avg)
```

Now we plot it
```{r}
library(viridis)
library(ggTimeSeries)



calendar.plot <- ggplot_calendar_heatmap(calendar.data,
                                         'Date',
                                         'Crimes') 
calendar.plot +
  xlab(NULL) +
  ylab(NULL) +
  scale_fill_viridis(alpha = 0.7)

```


```{r}
# creating some data


library(ggTimeSeries)
set.seed(1)
dtData = data.table(
      DateCol = seq(
         as.Date("1/01/2014", "%d/%m/%Y"),
         as.Date("31/12/2015", "%d/%m/%Y"),
         "days"
      ),
      ValueCol = runif(730)
   )
dtData[, ValueCol := ValueCol + (strftime(DateCol,"%u") %in% c(6,7) * runif(1) * 0.75), .I]
dtData[, ValueCol := ValueCol + (abs(as.numeric(strftime(DateCol,"%m")) - 6.5)) * runif(1) * 0.75, .I]

# base plot
p1 = ggplot_calendar_heatmap(
   dtData,
   'DateCol',
   'ValueCol'
)

# adding some formatting
p1 + 
   xlab(NULL) + 
   ylab(NULL) + 
   scale_fill_continuous(low = 'green', high = 'red') + 
   facet_wrap(~Year, ncol = 1)
```


