---
title: "Fargo_Crime_2017"
output: html_notebook
---

```{r, warning = F}
library(tidyverse)
library(ggmap)
library(lubridate)
library(ggthemes)
library(scales)
```


#### Upload Data

Try uploading test of the full data, not all lat/lon included
```{r}
full <- read.csv("crime.csv", header = T)
```

```{r}
str(full)
```
#### Data Wrangling

Convert "Date.Time" field into actual date field
```{r}
full$Date.Time <- lubridate::mdy(full$Date.Time)
```


Convert "Time" field into actual time field
```{r}
library(lubridate)

full$Time <- lubridate::mdy_hms(full$Time)

```


```{r}
str(full)
```


Drop the NA rows
```{r}
full <- tidyr::drop_na(data = full)
```

```{r}
str(full)
```


```{r}
head(full)
```

Recode the Call.Type to smaller groups
```{r}
full.recode <- dplyr::mutate(full, Category = recode(full$Call.Type,   # make new categories
                                                       "Abuse" = "Violent",
                                                       "Assault" = "Violent",
                                                       "Bicycle - Theft" = "Property",
                                                       "Bomb Threat" = "Violent",
                                                       "Burglary" = "Property",
                                                       "Death Notification" = "Violent",
                                                       "Domestic" = "Violent",
                                                       "Graffiti" = "Property",
                                                       "Gun Shots Heard" = "Violent",
                                                       "Harassment" = "Violent",
                                                       "Impaired Driver" = "Drugs/Alcohol",
                                                       "Impaired Person" = "Drugs/Alcohol",
                                                       "Kidnapping" = "Violent",
                                                       "Loud Party/People" = "Drugs/Alcohol",
                                                       "Overdose - Poisoning" = "Drugs/Alcohol",
                                                       "Prowler/Peeper" = "Violent",
                                                       "Robbery" = "Property", 
                                                       "Sexual Assault" = "Violent",
                                                       "Stab/Gunshot/Penetrating" = "Violent",
                                                       "Stolen Vehicle" = "Property",
                                                       "Theft" = "Property",
                                                       "Vandalism" = "Property",
                                                       "Vehicle Break In" = "Property"))


    ```



Mutate a new column with the hour the crime took place within
```{r}
full.spread <- dplyr::mutate(full, Time.Hour = hour(full$Time))
```




## 1. Crime category by time of day
```{r}

ggplot(full.recode, aes(x = Time)) + 
  geom_density(aes(fill = as.factor(Category),
                   colour = Category, y = ..density..),
               stat = "density",
               position = "identity",
               alpha = 0.4) + 
  theme_economist(base_size = 10) + 
  theme(text = element_text(family = "sans",
                          face = "plain",
                          color = "#000000",
                          size = 8,
                          hjust = 0.5,
                          vjust = 0.5)) + 
  guides(fill = guide_legend(title = "Category"), 
         colour = guide_legend(title = "Category"))  + 
  ylab("density") + 
  xlab("Time of Day") + 
  ggtitle("") +
  scale_x_datetime(labels = date_format("%H:%M")) + 
  theme(axis.text.y = element_blank()) 
scale_x_datetime()
ggsave(filename = "plot1.png")
```










Plot Calendar Heatmap
using ggTimeSeries

Wrangle data into occurences by day
```{r}
library(data.table)
calendar.data <- dplyr::add_count(full.recode, Date.Time)
calendar.data <- dplyr::arrange(calendar.data, Date.Time)

calendar.data <- calendar.data %>% 
  group_by(Date.Time) %>% 
  summarise(avg = mean(n))

# need to coerce to data.table
calendar.data <- data.table::data.table(Date = calendar.data$Date.Time,
                                        Crimes = calendar.data$avg)
```

Now we plot it
```{r}
library(viridis)
library(ggTimeSeries)



calendar.plot <- ggplot_calendar_heatmap(calendar.data,
                                         'Date',
                                         'Crimes') 
calendar.plot +
  xlab(NULL) +
  ylab(NULL) +
  scale_fill_viridis(alpha = 0.7)

```

```{r}
# order by month
full$Month <- factor(full$Month, c("Jan", "Feb", "Mar", "Apr", "May",
                                   "Jun", "Jul", "Aug", "Sep", "Oct",
                                   "Nov", "Dec"))

ggplot(data = full, aes(Month)) +
  geom_bar(stat = "count", fill = "indianred1", alpha = 0.7) +
  ggthemes::theme_economist(base_size = 10) 

#ggplot2::ggsave(filename = "plot2.jpg")
```

```{r}
# order by weekday
full$Weekday <- factor(full$Weekday, c("Sunday", "Monday", "Tuesday", "Wednesday",
                                       "Thursday","Friday", "Saturday"))

ggplot(data = full, aes(Weekday)) +
  geom_bar(stat = "count", fill = "indianred1", alpha = 0.7) +
  ggthemes::theme_economist(base_size = 10) 

#ggplot2::ggsave(filename = "plot3.jpg")
```







Where is the most crime?

```{r}
library(leaflet)
library(leaflet.extras)
```

```{r}
leafplot <- leaflet::leaflet(data = full) %>%
  addProviderTiles(providers$OpenStreetMap.DE) %>%
  addCircles(lng = full$Lon,
             lat = full$Lat)

leafplot
htmlwidgets::saveWidget(widget = leafplot,
                        file = "leafplot.html",
                        selfcontained = T)
```


What about using ggplot and gganimate?


```{r}
library(gganimate)
```

```{r}
full.recode.order <- full.recode
full.recode.order$Month <- recode(full.recode.order$Month,'Jan' = 1, 'Feb' = 2, 'Mar' = 3, 'Apr' = 4,
                                          'May' = 5, 'Jun' = 6, 'Jul' = 7, 'Aug' = 8,
                                          'Sep' = 9, 'Oct' = 10, 'Nov' = 11, 'Dec' = 12)
```


```{r}
# order by month
full.recode$Month <- factor(full.recode$Month, c("Jan", "Feb", "Mar", "Apr", "May",
                                   "Jun", "Jul", "Aug", "Sep", "Oct",
                                   "Nov", "Dec"))


myLocation_OD <- c(lon = median(full.recode$Lon),
                   lat = median(full.recode$Lat))

myMAP_OD <- get_map(myLocation_OD,
                    source = "google",
                    maptype = "roadmap",
                    zoom = 12)
myMAP_OD
```

Pointmap
```{r}


plotpoints_OD <- ggmap(myMAP_OD, darken = c(0.1, "white"), extent = "device") + geom_point(data = full.recode.order,
                                        aes(frame = full.recode.order$Month, cumulative = TRUE, x = Lon, y = Lat, color = Category),
                                        size = 3 ,
                                        alpha = 0.4) +
  ggtitle("Crime Growth in 2017") +
  theme(legend.position = c(0.83, 0.16)) 
  


gganimate(plotpoints_OD, filename = "animatedcrime.gif",interval = 1.0, title_frame = F)
#gganimate(plotpoints_OD, "animatedcrime.gif")

```


