---
title: "Fargo_Crime_2017"
output: html_notebook
---

```{r, warning = F}
library(tidyverse)
library(ggmap)
library(lubridate)
```


```{r}
my_OD <- read.csv("Fargo_OD.csv", header = T)
```

```{r}
str(my_OD)
```

Plot it   
Base Map
```{r}
myLocation_OD <- c(lon=median(my_OD$Lon),
                   lat=median(my_OD$Lat))

myMAP_OD <- get_map(myLocation_OD,
                    source = "google",
                    maptype = "terrain",
                    zoom = 12)
```

Pointmap
```{r}
plotpoints_OD <- ggmap(myMAP_OD, darken = c(0.3, "white")) + geom_point(data=my_OD,
                                        aes(x=Lon, y=Lat),
                                        size=2,
                                        color="red",
                                        alpha=.8)
plotpoints_OD
```

Now plot density
```{r}
plotdensity_OD <- ggmap(myMAP_OD, darken = c(0.0, "white")) +
                         stat_density2d(aes(x = Lon, y = Lat, fill = -..level.., alpha = ..level..),
                                        size = 1,
                                        bins = 15,
                                        data = my_OD,
                                        geom = "polygon") +
                         scale_fill_gradientn(colors = rainbow(4)) +
                         scale_alpha(range = c(0.1, 0.1),
                                     guide = F) +
                         ggtitle("2017 Fargo Overdose Calls Heatmap") +
                         labs(subtitle = "Brad Lindblad, MBA",
                              caption= "Source: http://fargond.gov/city-government/departments \n /police/police-records-data/dispatch-logs") +
                         theme(legend.position="none",
                               plot.subtitle = element_text(color="#666666"),
                               plot.caption = element_text(color="#AAAAAA", size=6)) +
                         ggsave("2017 Fargo Overdose Calls Heatmap.png") 



plotdensity_OD
```

Zoom in downtown
```{r}
myLocation_OD_zoom <- c(lon=-96.79,
                lat=46.876)

myMAP_OD_zoom <- get_map(myLocation_OD_zoom,
                 maptype = "hybrid",
                 zoom = 15)

myMAP_OD_zoom <- 
plotdensity_OD_zoom <- ggmap(myMAP_OD_zoom, darken = c(0.0, "white")) +
                       stat_density2d(
                       aes(x = Lon, y = Lat, fill = -..level.., alpha = ..level..),
                       size = 2, bins = 12, data = my_OD, geom = "polygon") +
                       scale_fill_gradientn(colours = rainbow(3)) +
                       scale_alpha(range = c(0.1, 0.2), guide = F) +
                       ggtitle("2017 Fargo Overdose Calls Heatmap") +
                       ggsave("2017 Fargo Overdose Calls Heatmap.png") +
                       labs(subtitle = "Source = Brad Lindblad, MBA") +
                       theme(legend.position="none")



plotdensity_OD_zoom
```


Try uploading test of the full data, not all lat/lon included
```{r}
full <- read.csv("crime.csv", header = T)
```

```{r}
str(full)
```

Convert "Date.Time" field into actual date field
```{r}
full$Date.Time <- lubridate::mdy(full$Date.Time)
```


Convert "Time" field into actual time field
```{r}
library(lubridate)

full$Time <- lubridate::mdy_hms(full$Time)

```


```{r}
str(full)
```


Drop the NA rows
```{r}
full <- tidyr::drop_na(data = full)
```

```{r}
str(full)
```


```{r}
head(full)
```

```{r}
ggplot(data = full) +
  geom_bar(mapping = aes(full$Month)) 
```
Now, build our first plotting function.   
This one called plotCrime, will simply plot the crime type in a heatmap over Fargo.
```{r}
plotCrime_test <- function(title = "title" ){
  # x = The main title for the graph,
  # Note center of Fargo using the median latitude and longitude
  myLocation_OD <- c(lon=median(my_OD$Lon),
                     lat=median(my_OD$Lat))
  # Build basemap
  myMAP_OD <- get_map(myLocation_OD,
                      source = "google",
                      maptype = "terrain",
                      zoom = 12)
  
  # Plot actual map
  ggmap(myMAP_OD, darken = c(0.0, "white")) +
                         stat_density2d(aes(x = Lon, y = Lat, fill = -..level.., alpha = ..level..),
                                        size = 1,
                                        bins = 15,
                                        data = my_OD,
                                        geom = "polygon") +
                         scale_fill_gradientn(colors = heat.colors(4)) +
                         scale_alpha(range = c(0.1, 0.1),
                                     guide = F) +
                         ggtitle(title) +
                         labs(subtitle = "Brad Lindblad, MBA",
                              caption="Data: http://fargond.gov/city-government/departments \n /police/police-records-data/dispatch-logs") +
                         theme(legend.position="none",
                               plot.subtitle = element_text(color="#666666"),
                               plot.caption = element_text(color="#AAAAAA", size = 6),
                               plot.title = element_text(color = "#666666", size = 25)) 
                        # ggsave(paste(title),'.png') 
  
}
```


Next function plots crime but allows us to select the crime type with a subset.  
```{r}
plotCrime_old <- function(title = "title", type = "type"){
  # title = The main title for the graph, type = type of crime to filter on
  
  
  # Pull data
  mydata <- full  # this is our cleaned data
  # Note center of Fargo using the median latitude and longitude
  myLocation_OD <- c(lon=median(mydata$Lon),
                     lat=median(mydata$Lat))
  # Build basemap
  myMAP_OD <- get_map(myLocation_OD,
                      source = "google",
                      maptype = "hybrid",
                      zoom = 12)

  # Subset the data
  mydata <- dplyr::filter(mydata, Call.Type == type)
  
  
  # Plot actual map
  ggmap(myMAP_OD, darken = c(0.0, "white")) +
                         stat_density2d(aes(x = Lon, y = Lat, fill = -..level.., alpha = ..level..),
                                        size = 1,
                                        bins = 15,
                                        data = mydata,
                                        geom = "polygon") +
                         scale_fill_gradientn(colors = topo.colors(4)) +
                         scale_alpha(range = c(0.1, 0.2),
                                     guide = F) +
                         ggtitle(title) +
                         labs(subtitle = "",
                              caption="Data: http://fargond.gov/city-government/departments \n /police/police-records-data/dispatch-logs") +
                         theme(legend.position="none",
                               plot.subtitle = element_text(color="#666666"),
                               plot.caption = element_text(color="#AAAAAA", size = 6),
                               plot.title = element_text(color = "#666666", size = 25),
                               axis.text.x = element_text(angle = 90, hjust = 1))
                        # ggsave(paste(title),'.png') 
  
  
}
```



Try plotting just points for a "heatmap"
```{r}
plotCrime_1 <- function(title = "title", type = "type"){
  # title = The main title for the graph, type = type of crime to filter on
  
  
  # Pull data
  mydata <- full  # this is our cleaned data
  # Note center of Fargo using the median latitude and longitude
  myLocation_OD <- c(lon=median(mydata$Lon),
                     lat=median(mydata$Lat))
  # Build basemap
  myMAP_OD <- get_map(myLocation_OD,
                      source = "google",
                      maptype = "hybrid",
                      zoom = 12)

  # Subset the data
  mydata <- dplyr::filter(mydata, Call.Type == type)
  
  
  # Plot actual map
  ggmap(myMAP_OD, darken = c(0.0, "white")) +
                         geom_point(data=my_OD,
                                        aes(x=Lon, y=Lat),
                                        size=5,
                                        color="red",
                                        alpha=0.3) +
                         ggtitle(title) +
                         labs(subtitle = "",
                              caption="Data: http://fargond.gov/city-government/departments \n /police/police-records-data/dispatch-logs") +
                         theme(legend.position="none",
                               plot.subtitle = element_text(color="#666666"),
                               plot.caption = element_text(color="#AAAAAA", size = 6),
                               plot.title = element_text(color = "#666666", size = 25),
                               axis.text.x = element_text(angle = 90, hjust = 1))
                        # ggsave(paste(title),'.png') 
  
  
}
```

Now, we'll make the heatmap fn for the shiny app:
```{r}
plotCrime <- function(title = "title", type = "type"){
  # title = The main title for the graph 
  # type = type of crime to filter on
 
  # Read data and transform
  full <- read.csv("crime.csv", header = T)
  full$Date.Time <- lubridate::mdy(full$Date.Time)
  full$Time <- lubridate::mdy_hms(full$Time)
 
  # Note center of Fargo using the median latitude and longitude
  myLocation_OD <- c(lon=median(full$Lon),
                     lat=median(full$Lat))
  # Build basemap
  myMAP_OD <- get_map(myLocation_OD,
                      source = "google",
                      maptype = "hybrid",
                      zoom = 12)
                    

  # Subset the data
  mydata <- dplyr::filter(full, Call.Type == type)
  
  
  # Plot actual map
  ggmap(myMAP_OD, darken = c(0.0, "white")) +
                         stat_density2d(aes(x = Lon, y = Lat, fill = -..level.., alpha = ..level..),
                                        size = 1,
                                        bins = 15,
                                        data = mydata,
                                        geom = "polygon") +
                         scale_fill_gradientn(colors = topo.colors(4)) +
                         scale_alpha(range = c(0.1, 0.2),
                                     guide = F) +
                         ggtitle(title) +
                         labs(subtitle = "",
                              caption="Data: http://fargond.gov/city-government/departments \n /police/police-records-data/dispatch-logs") +
                         theme(legend.position="none",
                               plot.subtitle = element_text(color="#666666"),
                               plot.caption = element_text(color="#AAAAAA", size = 6),
                               plot.title = element_text(color = "#666666", size = 25),
                               axis.text.x = element_text(angle = 90, hjust = 1))  +
                         coord_fixed(xlim = c(-96.93,-96.74), ylim = c(46.79, 46.935), ratio = 1/1)
                         
                        # ggsave(paste(title),'.png') 
  
  
}
```

```{r}
df <- data.frame(
     x = runif(100, 0, 5),
     y = runif(100, 0, 5))
```

```{r}
library(ggplot2)
ggplot(df, aes(x=x, y=y)) + geom_point()
```

```{r}
ggplot(df, aes(x=x, y=y)) + geom_point() + coord_fixed(0.5)
```

